Generate Mortality Adjustment for FEM

### 2021-01-20 FEM_R/mortality_adjustment.Rmd
Updated the mortality adjustment to use a different data source. Now using mortality rates (qx) to calculate the adjustment, which is much more similar to the US values now

require(tidyverse)
require(readODS)

@@

>>>>

require(tidyverse)
require(readODS)
require(ggplot2)

@@

# DISREGARD ALL THE ABOVE

It's all wrong. Start again but can reuse some of the code for another dataset.

# Second Attempt Using Mortality Rates (qx)

```{r}
m_period <- read_ods('../Mortality_Projections/engppp18qx.ods',
                     sheet = 3,
                     skip = 6,
                     row_names = TRUE)
f_period <- read_ods('../Mortality_Projections/engppp18qx.ods',
                     sheet = 4,
                     skip = 6,
                     row_names = TRUE)
m_cohort <- read_ods('../Mortality_Projections/engppp18qx.ods',
                     sheet = 5,
                     skip = 6,
                     row_names = FALSE)
f_cohort <- read_ods('../Mortality_Projections/engppp18qx.ods',
                     sheet = 6,
                     skip = 6,
                     row_names = FALSE)

- Drop ages below 50
- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
- Calculate ratio between values for consecutive years

```{r}
# Rename age column and filter out ages below 50
male <- m_cohort %>% 
            rename(age = 'age (years)') %>%
            filter(age >= 50)

female <- f_cohort %>% 
            rename(age = 'age (years)') %>%
            filter(age >= 50)

ratio_df <- data.frame(matrix(vector(mode = 'numeric', length = nrow(male) * ncol(male)), nrow = nrow(male), ncol = ncol(male)))

for(i in nrow(male)) {
  for(j in ncol(male)) {
    ratio_df[i,j] <- male[i,j] / male[i-1, j-1]
  }
}

```{r}
# Rename age column and filter out ages below 50
male <- m_cohort %>% 
            rename(age = 'age (years)') %>%
            filter(age >= 50)

female <- f_cohort %>% 
            rename(age = 'age (years)') %>%
            filter(age >= 50)

# Scale all values to the 2012 column
# SET TO 2004 FOR TESTING AGAINST US, NEED TO CHANGE THIS BACK 
male_norm <- male / male[,24]
female_norm <- female / female[,24]

head(female_norm)
```

```{r}
# Copy dfs so we don't have to rerun other chunks
m_copy <- male_norm
f_copy <- female_norm

Write out the file to E_FEM_clean/Mortality_Projections/.

```{r}
# Test against US data
us_adj <- read_tsv('../Mortality_Projections/mortality_adj_US.txt', col_names = TRUE, skip = 1)

adj <- adjustment

adj$year <- as.numeric(as.character(adj$year))
us_adj$year <- as.numeric(as.character(us_adj$year))

# Extract only 60 year olds from my dataset
# Drop all years before 1990 as this is when the US data starts
adj_60 <- adj %>%
                filter(age == 60) %>%
                select(-age) %>%
                filter(year > 1989) %>%
                rename(group = male)

# Drop all years after 2068 for US as this is when Eng data stops
# Also drop over 65 values
adj_us <- us_adj %>%
                filter(year < 2069) %>%
                filter(l2age65pd == 0) %>%
                rename(group = l2age65pd)
adj_us$group <- 2

NOW PLOT!!

```{r}

# Now combine the data and start plotting
combined <- rbind(adj_60, adj_us)

combined$group[combined$group == 1] <- 'male'
combined$group[combined$group == 0] <- 'female'
combined$group[combined$group == 2] <- 'US'

ggplot(data = combined, aes(x = year, y = adjustment, color = group)) +
      geom_line(size = 0.5) +
      labs(title = 'Age 60')

cut <- combined %>% filter(group != 'US')

ggplot(data = cut, aes(x = year, y = adjustment, color = group)) +
      geom_line()

cut2 <- cut %>% filter(group == 'female')

ggplot(data = cut2, aes(x = year, y = adjustment, color = group)) +
      geom_line()

```{r}
# Test against US data
us_adj2 <- read_tsv('../Mortality_Projections/mortality_adj_US.txt', col_names = TRUE, skip = 1)

adj2 <- adjustment

adj2$year <- as.numeric(as.character(adj2$year))
us_adj$year <- as.numeric(as.character(us_adj$year))
# Extract only 60 year olds from my dataset
# Drop all years before 1990 as this is when the US data starts
adj_60 <- adj2 %>%
                filter(age == 75) %>%
                select(-age) %>%
                filter(year > 1989) %>%
                rename(group = male)

# Drop all years after 2068 for US as this is when Eng data stops
# Also drop over 65 values
adj_us <- us_adj %>%
                filter(year < 2069) %>%
                filter(l2age65pd == 1) %>%
                rename(group = l2age65pd)
adj_us$group <- 2

```
```{r}

# Now combine the data and start plotting
combined <- rbind(adj_60, adj_us)

combined$group[combined$group == 1] <- 'male'
combined$group[combined$group == 0] <- 'female'
combined$group[combined$group == 2] <- 'US'

ggplot(data = combined, aes(x = year, y = adjustment, color = group)) +
      geom_line(size = 0.5) +
      labs(title = 'Age 75')

cut <- combined %>% filter(group != 'US')

ggplot(data = cut, aes(x = year, y = adjustment, color = group)) +
      geom_line()

cut2 <- cut %>% filter(group == 'female')

ggplot(data = cut2, aes(x = year, y = adjustment, color = group)) +
      geom_line()
